Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-08 paper.tex
re- factored experiments

\section{Experiments}

\subsection{Boolean circuits}
As a toy domain, we consider the problem of learning
to build the Boolean circuits out of logic gates.
Now, if we equipped our agent with a full repertoire of
circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
Instead, we gave \system only a NAND gate,
and tasked it with building a set of 250 random circuits made out of
AND, NOT, and OR gates,
as well as 2 or 4 input multiplexers.
\system observes the input/output truth table of the circuits.
This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
%% For this domain our recognition model is a simple multilayer perceptron 
%% whose inputs are the input/output examples.
\system learns to build these
circuits by inventing new subroutines that correspond to
logic gates and multiplexers (Figure~\ref{multiplexer}, top).
We compare  with four baselines on 250 held-out circuits:
No NN, which lesions $q$;
RF/DC,
which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a $q$ only from samples from the fixed generative model --
this is how RobustFill and DeepCoder are trained;
PCFG, which lesions $q$, fixes $\mathcal{D}$, and learns only $\theta$ (so it is analogous to learning the parameters of a PCFG);
and Enum, which does no learning and just enumerates a large frontier.
\begin{figure}
\begin{tabular}{lllll}
  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
  84\% solved&76\%& 63\%&62\%&62\%\\
  $0.9$ sec&  1.1&1.2&1.3&1.3
  \\\bottomrule
\end{tabular}
\caption{Top: \% held-out Boolean circuits solved by \system and baselines w/ 5sec timeout. Bottom: average search time.}
\end{figure}

@@

    (x.out) -- (xz.in 1)
    (z.out) -- (xz.in 2)

@@

\begin{table}
\begin{tabular}{lllll}
  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
  98\% solved&94\%&0.7\%&0\%&2\% \\
  $2.7$ sec& $2.8$ sec  &$3.6$ sec&--&$2.1$ sec\\
  $9.3$ nats &$11$ nats&$2.3$ nats&--&$4.6$ nats
  \\\bottomrule
\end{tabular}\caption{Symbolic regression. Top: \% held out tasks solved w/ 5 second timeout. Middle: average solve time. Bottom: average description length $-\log \probability[x|p]$. ``No NN'' solves almost as many tasks but
using worse programs w/ extraneous degrees of freedom, increasing  description length.}\label{regressionComparison}
\end{table}

@@

\begin{table}
\begin{tabular}{lllll}
  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
  99\% solved\\
  $0.88\pm 0.04$ sec
  \\\bottomrule
\end{tabular}\caption{\% held out tasks solved and average solve time for string edit problems}
\end{table}

@@

\pagebreak
 \subsection{Related Work}
 Our work is far from the first to learn to learn programs,
 an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:

@@

 \section{Conclusion}
 We have contributed a model, \system, that learns to program by
 bootstrapping a domain-specific language for expressing the solutions
 to the tasks at hand.  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the desired programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.

>>>>

\section{Experiments}
We first present our experimental results qualitatively,
illustrating the DSLs that  \system
learns in the domains of Boolean circuits, symbolic regression, string editing, and list functions.
We then show quantitative results (Section~\ref{quantitative})
comparing \system with lesioned versions
of our model that are analogous to prior work.

\subsection{Boolean circuits}
As a toy domain, we consider the problem of learning
to build the Boolean circuits out of logic gates.
If we gave our agent the full repertoire of
circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
Instead, we gave \system only a NAND gate,
and tasked it with building a set of 250 random circuits made out of
AND, NOT, and OR gates,
as well as 2 or 4 input multiplexers.
\system observes the input/output truth table of the circuits.
This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
%% For this domain our recognition model is a simple multilayer perceptron 
%% whose inputs are the input/output examples.
\system learns to build these
circuits by inventing new subroutines that correspond to
logic gates and multiplexers (Figure~\ref{multiplexer}, top).

@@

    (x.out) -| (xz.in 1)
    (z.out) -- (xz.in 2)

@@

@@

@@

\subsection{Quantitative Results}\label{quantitative}

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours--NN}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (see Section~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Section~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{PCFG}, which lesions the recognition model, fixes $\mathcal{D}$, and learns only $\theta$.
This is like learning the parameters of a PCFG.
\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
This is equivalent to the first wake cycle of our \system.

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly can find those solutions.
For symbolic regression, we also care about the quality
of the solution, as measured by the likelihood model $\probability[x|p]$.
Concretely, this cashes out as measuring how many
real numbers were used in the programs: did the model correctly explain a linear function
using two numbers, or did it introduce extraneous parameters?
Table~\ref{baselineComparisons}
compares our model against these baselines.

\begin{table}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours&Ours--NN  & RF/DC & PCFG & Enum

  \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
  \% solved &
  84\% &76\%& 63\%&62\%&62\%\\
  Solve time&  $0.9$s&  1.1&1.2&1.3&1.3

  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&    98\% &94\%&0.7\%&0\%&2\% \\
  Solve time&  $2.7$s& $2.8$s  &$3.6$s&--&$2.1$s\\
  MDL (nats)&   $9.3$ &$11.0$&$2.3$&--&$4.6$

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% \\
  Solve time&  0.9s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}
\pagebreak
 \section{Related Work}
 Our work is far from the first to learn to learn programs,
 an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:

@@

 \section{Conclusion}
 We contributed a model, \system, that learns to program by
 bootstrapping a domain-specific language.
 The  DSL learns to expose new domain-specific primitives
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the desired programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.
